---
title: "Recommended Print List"
author: "Ken Brooks"
date: "November 13, 2015"
output: pdf_document
---

Prepare Environment
-------------------

```{r libraries, warning=FALSE, echo=FALSE, error=FALSE}
# clear environment and load libraries
rm(list = ls())

# note: use install_packages.Rmd for installation
library(xts)
library(readxl)
library(plyr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(lubridate)
library(forecast)
source("Inv_Planning.R")


```

Import data
-----------

```{r read, echo=FALSE}

filename <- "Sales by month v3.csv"

data = read.csv(filename)
```

Prepare data
------------

```{r format.data}

checkData(data)  # check that nothing has changed in input format
data <- cleanImportedData(data)  # cleanup imported data
saveRDS(data, "data.rds")

```

Build data frames
-----------------

```{r data.frames}

# Retrieve stored data
data = readRDS("data.rds")

title.data <- createTitleDF(data)  # create title metadata df
time.series.data <- createTSDF(data)  # create time.series df


saveRDS(title.data, "title_data.rds")
saveRDS(time.series.data, "time_series.rds")

```


Generate a single forecast
--------------------------

```{r}

source("Inv_Planning.R")
time.series.data = readRDS("time_series.rds")

isbn = "0716756676"

fcst <- n.month.forecast(isbn, time.series.data, 15)

fcst


```



Forecast all titles
-------------------

```{r forecast, warning=FALSE, error=FALSE}
# The purpose of this chunk is to generate forecasts for many titles at once

# Get all of the time series and title data


title.forecasts <- build.forecast.DF(title.data, time.series.data)

# Check the csv for strings in the ISBNs - dropping leading zeros
# replace all negative forecasts with 0 before doing the sum (or do this
# on the base forecast?)

```

Align previous edition with current and plot
--------------------------------------------

```{r align.previous.edition}
# Make current edition and previous edition start with the same year.

time.series.data = readRDS("time_series.rds")
title.data = readRDS("title_data.rds")

source("Inv_Planning.R")


isbn1 = "1319056911"  # How Life Works
isbn2 = "0312132921"
isbn3 <- "031243829X"



# Test plot.previous.edition
align.previous.edition(isbn1, time.series.data, title.data)
# align.previous.edition(isbn2, time.series, title.data)

isbn3 <- "031243829X"
align.previous.edition(isbn3, time.series.data, title.data)  # should fail

```

Output multiple forecasts
-------------------------

```{r output.multiple, warning=FALSE, error=FALSE}
# The purpose of this chunk is to generate many different forecasts for 
# a single title

# Get all of the time series and title data
time.series = readRDS("time_series.rds")
title.data = readRDS("title_data.rds")

# Function to return a forecast of 0s:

forecast.zeros <- function(ts, horizon) {
  # takes a time series and a horizon and returns 
  # a 0 forecast over that horizon
  ts.temp = ts(rep(0,length(ts)), 
                start = start(ts), 
                frequency = frequency(ts))
    ts.fcst = forecast(ts.temp, h=horizon)
    return(ts.fcst)
  }

# create a function to forecast a series based on a ts and df index

output.multiple <- function(isbn){  # note that this should be a 2 param fn
  
  # Generates multiple versions of a forecast to a csv file.  Takes as input 
  # ISBN and implicit time.series and title.data data frames
  
  horizon = 15
  isbn.prev <- previous.edition.isbn(isbn, title.data) # Get previous edition
  ts1 = time.series[, isbn]
  
  num.contig <- length(na.contiguous(ts1))  # how many contiguous values?
  
  if (num.contig > 6){  # if there's enough data, generate a forecast
    
    # base forecast
    base.forecast = forecast(ts1, h=horizon)
    
  } else {   # if there's not enough data, generate a zero forecast
    
    # zeroes forecast
    base.forecast = forecast.zeros(ts1, horizon)
    plot(base.forecast)

  }
  
  plot(base.forecast)
  output <- base.forecast$mean

  # If there's a previous edition, generate additional forecasts

  if (!is.na(isbn.prev)) {
    
    # generate a forecast from current edition plus previous edition
    
    ts.summed <- sum.edition.data(isbn, time.series, title.data)
    summed.prev.fcst = forecast(ts.summed, h = horizon)
    plot(summed.prev.fcst)
    output <- cbind(output, summed.prev.fcst$mean)
    
    # generate a forecast that is simply the previous editions data
    
    prev.edition <- align.previous.edition(isbn, time.series, title.data) # aligns data
    
    # take the window associated with the forecast
    
    prev.edition.2 <- window(prev.edition, 
                      start= start(summed.prev.fcst$mean),
                      end = end(summed.prev.fcst$mean),
                      extend = F)

    output <- cbind(output, prev.edition.2)
    
    }
  
  # print(output)
  
  # note that the date formatting needs to be fixed
  # write.csv(file=paste(isbn, ".csv", sep = ""), x=as.data.frame(output))
  write.zoo(output,file=paste(isbn, ".csv", sep = ""), index.name="Date",sep=",")
}

isbn = "1319056911"  # How Life Works
output.multiple(isbn)

# isbn2 = "0312132921"
# output.multiple(isbn2)
# isbn3 <- "031243829X"
# output.multiple(isbn3)

```


Review Individual Titles
------------------------

```{r}

time.series = readRDS("time_series.rds")
title.data = readRDS("title_data.rds")

title.data$Isbn

isbn1 <- "0312132921"
isbn2 <- "031241501X" # skewed title 1
isbn3 <- "312566735" # skewed title 2
isbn4 <- "145760647X" # skewed title 3

skewed.isbns <- as.list(c(isbn2, isbn3, isbn4))
ts1 = time.series[, isbn4]
sum(!is.na(ts1))
ts1.fcst = forecast(ts1, h = 15)
plot(ts1.fcst)

# test actions.
str(ts1.fcst)
ts1.fcst$method
print(length(na.contiguous(ts1)))


```

Plot two time series on a single axis
-------------------------------------

```{r plot.two}

isbn1 <- "031241501X"
isbn2 <- "031243829X"
time.series = readRDS("time_series.rds")

plot.two <- function(isbn1, isbn2, ts) {
  # Plots two ISBNs in a facet plot, one over the other.  Takes 2 ISBNs as
  # strings and a time series data frome as input and generates two plots.
  plot.data <- ts[,c(isbn1, isbn2)]
  plot.data <- zoo::na.trim(plot.data, sides = "left", is.na = "all")
  plot(plot.data, main = "Comparative Time Series Plot", 
       xlab = "Date",
       ylab = "Units",
       plot.type = "s", 
       col = 1:2)
  legend("topright", 
         legend = c(isbn1, isbn2),
         lty = c(1,1),
         col = 1:2)
  }

plot.two(isbn1, isbn2, time.series)


```

Plot title and previous edition (if any)
----------------------------------------

```{r previous.edition}

time.series = readRDS("time_series.rds")
title.data = readRDS("title_data.rds")

isbn1 = "1319056911"  # How Life Works
isbn2 = "0312132921"
isbn3 <- "031243829X"

plot.previous.edition <- function(isbn, ts.data, title.data){
  # Takes an ISBN as a string and two data frames: title data and title
  # time series as inputs and plots the time series and
  # the previous edition.  If there is no previous edition, it returns 
  # an NA.
  
  isbn.prev <- previous.edition.isbn(isbn, title.data) # Get previous edition
  
  # Fail if there's no previous edition (isbn.prev == NA)
  if (is.na(isbn.prev)) {
    return(NA)
    }
  
  # Fail if previous edition has no data (not yet implemented)
  
  # Plot results
  plot.two(isbn1, isbn.prev, time.series)
}

# Test plot.previous.edition
plot.previous.edition(isbn1, time.series, title.data)
plot.previous.edition(isbn2, time.series, title.data)

isbn3 <- "031243829X"
plot.previous.edition(isbn3, time.series, title.data)

```




Test decomposition with STL

```{r}

isbn = "1319056911"  # How Life Works
isbn.fit = stl(isbn, s.window = "periodic")
plot(isbn.fit)

```

Transplant model from previous edition
--------------------------------------
* Add the previous edition to the current and forecast that
* Decompose the previous edition and use the seasonal factors to forecast the current edition, based on current trend

```{r test.transplant.forecast}

# Make current edition and previous edition start with the same year.

time.series = readRDS("time_series.rds")
title.data = readRDS("title_data.rds")

isbn = "1319056911"  # How Life Works
isbn2 = "0312132921"
isbn3 <- "031243829X"

forecast.with.previous <- function(isbn, ts.data, title.data){
  # Takes an ISBN as a string and two data frames: title data and title
  # time series as inputs and uses the model from the previous edition
  # to forecast the new edition. Returns an NA if there's no previous
  # edition.
  
  isbn.prev <- previous.edition.isbn(isbn, title.data) # Get previous edition
  
  # Fail if there's no previous edition (isbn.prev == NA)
  if (is.na(isbn.prev)) {
    return(NA)
    }
  # Fail if previous edition has no data (not yet implemented)
  ts1 = time.series[, isbn]
  ts1 <- zoo::na.trim(ts1, sides = "both", is.na = "all")
  print(ts1)
  ts1.fcst = forecast(ts1, h=15)
  plot(ts1.fcst)
  
  ts2 = time.series[, isbn.prev]
  ts2 <- zoo::na.trim(ts2, sides = "both", is.na = "all")
  print(ts2)
  ts2.fcst = forecast(ts2, h=15)
  plot(ts2.fcst)
  
  temp = merge.zoo(ts1, ts2)
  temp[is.na(temp)] <- 0
  ts3 = temp[,1] + temp[,2]
  ts3 = ts(ts3, start = start(ts3), frequency = 12)
  ts3[ts3 < 0] <- 0
  print(ts3)
  ts3.fcst = forecast(ts3, h=15)
  plot(ts3.fcst)

}

# Test plot.previous.edition
forecast.with.previous(isbn1, time.series, title.data)
# align.previous.edition(isbn2, time.series, title.data)

isbn3 <- "031243829X"
forecast.with.previous(isbn3, time.series, title.data)  # should fail

```

Recommend titles to examine
---------------------------

```{r}

title.data = readRDS("title_data.rds")
title.forecasts = readRDS("forecasts.rds")

# add another column for whether a review is required

calc.review <- function(forecast, onhand){
  if(forecast == 0){"Insuff data"}
  else if(forecast > onhand){"Print review"}
  else{"No Review"}
}

title.planning <- title.data %>% rowwise() %>% 
  dplyr::mutate(Review = calc.review(title.forecasts$three, (Available + Reserved.Stock + Sets.Made.Up))) %>% 
  dplyr::arrange(Review) %>% 
  mutate(OQ.6 = title.forecasts$six) %>% 
  mutate(OQ.9 = title.forecasts$nine) %>% 
  mutate(OQ.12 = title.forecasts$twelve) %>% 
  mutate(OQ.15 = title.forecasts$fifteen)

# output to csv

saveRDS(title.planning, "title_planning.rds")
write.csv(file="Title_Planning.csv", x=title.planning)

```

Release Notes
-------------
v6
* plot.two function
  * Eliminate leading blanks in plot.two function
  * Moved both lines into same panel for easier comparison
* align.previous edition
  * added function align.previous.edition to produce plot
  * it aligns peaks, not months - this may need to be changed
  * It probablly also needs an export, or to at least to print the prev isbn in the legend
v7
* skipped
v8
* sum.previous.edition function created
* output.multiple function created
