---
title: "Recommended Print List"
author: "Ken Brooks"
date: "November 13, 2015"
output: pdf_document
---

Prepare Environment
-------------------

```{r libraries, warning=FALSE, echo=FALSE, error=FALSE}
# clear environment and load libraries
rm(list = ls())

# note: use install_packages.Rmd for installation
library(xts)
library(readxl)
library(plyr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(lubridate)
library(forecast)


```

Import data
-----------

```{r read, echo=FALSE}
setwd("~/Documents/Macmillan/Projects/Demand Planning")

# filename <- file.choose()

# filename <- "Sales by month.csv"
filename <- "Sales by month v3.csv"

# data = read_csv(filename, sheet = 1, skip = 0)
data = read.csv(filename)
```

Check that nothing has changed in the input data
------------------------------------------------

```{r checkdata}

# str(data)
dim(data)[2] == 16
is.factor(data$Author)
is.factor(data$Isbn)
is.numeric(data$Ean)
is.factor(data$Title)
is.factor(data$Pub.Date)
is.factor(data$Medium)
is.factor(data$Ans.Code)
is.integer(data$Available)
is.integer(data$Consignment.On.Hand)
is.integer(data$Reserved.Stock)
is.integer(data$Sets.Made.Up)
is.integer(data$Month)
is.integer(data$Gross.Units)
is.integer(data$Return.Units)
is.factor(data$Prev.Edition.Isbn)
is.numeric(data$Prev.Edition.Ean)

```

Clean up imported data
----------------------

```{r format.data}

names(data)


valid_column_names <- make.names(names=names(data), unique=TRUE, allow_ = TRUE)
names(data) <- valid_column_names

data$Title <- gsub("..", ".", make.names(data$Title, allow_ = TRUE), fixed = TRUE)
data$Author <- gsub("..", ".", make.names(data$Author, allow = TRUE), fixed = TRUE)
data$Ean <- as.factor(as.character(data$Ean))
data$Prev.Edition.Ean <- as.factor(as.character(data$Prev.Edition.Ean))
data$Pub.Date <- as.Date(as.character(data$Pub.Date), format = "%m/%d/%Y")
# data$Month <- as.Date(as.character(data$Month), format = "%Y%m")

data$Month = as.Date(paste(as.character(data$Month), "01", sep = ""), format = "%Y%m%d")

# Eliminate duplicates, if any:
data <- unique(data)  # Note that this might cause problems later!

str(data)
summary(data)

saveRDS(data, "data.rds")

```

Build title table
-----------------

```{r title.table}

# Retrieve stored data
data = readRDS("data.rds")

names(data)

title.data = data %>% dplyr::select(Author, Isbn, Ean, Title, Pub.Date, Medium, Ans.Code, Available, Consignment.On.Hand, Reserved.Stock, Sets.Made.Up, Prev.Edition.Isbn, Prev.Edition.Ean) %>%
  distinct(Author, Isbn, Ean, Title, Pub.Date, Medium, Ans.Code, Available,
    Consignment.On.Hand, Reserved.Stock, Sets.Made.Up, Prev.Edition.Isbn, Prev.Edition.Ean)

saveRDS(title.data, "title_data.rds")

```

Build time series objects
-------------------------

```{r time.series, warning=FALSE, error=FALSE}
# The purpose of this chunk is to create time series objects for a large number of titles

# Retrieve stored data
data = readRDS("data.rds")

names(data)

demand = data %>% dplyr::select(Month, Isbn, Gross.Units) 

demand = tidyr::spread(demand, key = Isbn, value = Gross.Units)

# Get the earliest month in order to align dates

earliest.date = min(demand$Month)

rownames(demand) <- demand$Month
demand3 = demand[,-1] # Remove the Month column since it's now rownames

# Get the earliest month

demand4 = ts(demand3, start = c(year(earliest.date), month(earliest.date)), frequency = 12)

saveRDS(demand4, "time_series.rds")

```

Find previous ISBN
------------------

```{r prev.isbn}

title.data = readRDS("title_data.rds")

# head(title.data, 50)

isbn1 <- "1429231246"
isbn2 <- "0312132921"

previous.edition.isbn <- function(isbn, title.df){
  # Returns the previous edition ISBN.  Takes ISBN as a string, and a data frame
  # as an input and returns the previous ISBN as a string, or NA if no previous
  # ISBN is available.
  result <- as.character(title.df[title.df$Isbn == isbn, "Prev.Edition.Isbn"])
  if (result == "")
    {NA}
  else {result}
}


# Tests for previous.edition.isbn
previous.edition.isbn("1429231246", title.data) == "0716761211"
is.na(previous.edition.isbn("0312132921", title.data))

```

Create a combined title with previous edition data
--------------------------------------------------

```{r sum.edition.data}

sum.edition.data <- function(isbn, ts.data, title.data){
  # Takes an ISBN as a string and two data frames: title data and title
  # time series as inputs and returns a new time series with the summed
  # values from that ISBN and its previous edition. If there is no previous edition, 
  # it returns the data from the first edition only
  
  # Get ISBN data
  title.1 <- ts.data[,isbn]
  
  # Find previous edition
  isbn.prev <- previous.edition.isbn(isbn, title.data) # Get previous edition
  
  # If there's no previous edition return the first title's data

  if (is.na(isbn.prev)) {
    return(title.1)
    }
  # Fail if previous edition has no data (not yet implemented)
  
  # Get previous edition data
  title.2 <- ts.data[,isbn.prev]
  
  # Combine the two titles
  title.1 <- zoo::na.trim(title.1, sides = "both", is.na = "all") # trim NAs
  title.2 <- zoo::na.trim(title.2, sides = "both", is.na = "all") # trim NAs
  
  temp = merge.zoo(title.1, title.2)
  temp[is.na(temp)] <- 0
  title.3 = temp[,1] + temp[,2]
  title.3 = ts(title.3, start = start(title.3), frequency = 12)
  title.3[title.3 < 0] <- 0
  
  title.3
}

time.series = readRDS("time_series.rds")
title.data = readRDS("title_data.rds")

isbn1 = "1319056911"  # How Life Works
isbn2 = "0312132921"
isbn3 <- "031243829X"

# Test sum.edition.data
sum.edition.data(isbn1, time.series, title.data)

# Should match isbn3 data
isbn3 <- "031243829X"
print(time.series[,isbn3])
print(sum.edition.data(isbn3, time.series, title.data))
```

Forecast time series
--------------------

```{r forecast, warning=FALSE, error=FALSE}
# The purpose of this chunk is to generate forecasts for many titles at once

# Get all of the time series and title data
time.series = readRDS("time_series.rds")
title.data = readRDS("title_data.rds")

# create a function to forecast a series based on a ts and df index

n.mo.fcst2 <- function(isbn){  # note that this should be a 2 param fn
  ts1 = time.series[, isbn]
  # if there's more than 24 months of data, generate forecast,
  # else generate a zero forecast to drive a manual review.
  # This needs to generate the forecast to see if it's flat.
  # If it's flat, replace it with a zero forecast to drive
  # a manual review.
  if(sum(!is.na(ts1)) < 25){  
    ts1.temp = ts(rep(0,length(ts1)), start = start(ts1), frequency = frequency(ts1))
    ts1.fcst = forecast(ts1.temp, h=15)$mean
    }
  else{
    ts1.fcst = forecast(ts1, h=15)$mean}
  ts1.fcst
}

# use ldply to apply it to a list of ISBNs

title.forecasts <- ldply(title.data$Isbn, n.mo.fcst2)
# add names
title.data$Isbn
rownames(title.forecasts) <- as.character(title.data$Isbn)

rownames(title.forecasts[1,])

# Add cumulative forecast quantities. Note that titles are in rows!

title.forecasts <- title.forecasts %>% mutate(three = rowSums(.[1:3])) %>%
  mutate(six = rowSums(.[1:6])) %>%
  mutate(nine = rowSums(.[1:9])) %>%
  mutate(twelve = rowSums(.[1:12])) %>%
  mutate(fifteen = rowSums(.[1:15]))

saveRDS(title.forecasts, "forecasts.rds")
write.csv(file="Title_forecasts.csv", x=title.forecasts)

```

Align previous edition with current and plot
--------------------------------------------

```{r align.previous.edition}
# Make current edition and previous edition start with the same year.

time.series = readRDS("time_series.rds")
title.data = readRDS("title_data.rds")

isbn1 = "1319056911"  # How Life Works
isbn2 = "0312132921"
isbn3 <- "031243829X"

align.previous.edition <- function(isbn, ts.data, title.data){
  # Takes an ISBN as a string and two data frames: title data and title
  # time series as inputs and plots the time series and the previous
  # edition aligned at peak sales.  If there is no previous edition, 
  # it returns an NA.
  
  isbn.prev <- previous.edition.isbn(isbn, title.data) # Get previous edition
  
  # Fail if there's no previous edition (isbn.prev == NA)
  if (is.na(isbn.prev)) {
    return(NA)
    }
  # Fail if previous edition has no data (not yet implemented)
  
  # Find the peak sales month for each title
  title.1 <- ts.data[,isbn]
  title.1.max <- which.max(title.1)
  
  title.2 <- ts.data[,isbn.prev]
  title.2.max <- which.max(title.2)

  # Determine how much to shift the previous edition to match peaks. 
  # Shift should be in 12 month increments
  
  shift = title.1.max - title.2.max   #shifting by this amount would align peaks
  
  shift = 12 * round(shift / 12.0)   # Makes shifts in 12 month increments
  
  #make room in the title to allow for the shift
  title.3 <- window(title.2, 
                    start= start(title.2),
                    end = end(title.2) + shift,
                    extend = TRUE)
  
  title.3 <- lag(title.3, shift)  # do the shift
  
  # Trim leading and trailing NAs and plot
  plot.data <- cbind(title.1, title.3)
  plot.data <- zoo::na.trim(plot.data, sides = "both", is.na = "all")
  plot(plot.data, main = "Plot versus previous edition",
       xlab = "Date",
       ylab = "Units",
       plot.type = "s", 
       col = 1:2)
  legend("topright", 
         legend = c(isbn1, "previous"),
         lty = c(1,1),
         col = 1:2)
  
  return(title.3)
}

# Test plot.previous.edition
align.previous.edition(isbn1, time.series, title.data)
# align.previous.edition(isbn2, time.series, title.data)

isbn3 <- "031243829X"
align.previous.edition(isbn3, time.series, title.data)  # should fail

```

Output multiple forecasts
-------------------------

```{r output.multiple, warning=FALSE, error=FALSE}
# The purpose of this chunk is to generate many different forecasts for 
# a single title

# Get all of the time series and title data
time.series = readRDS("time_series.rds")
title.data = readRDS("title_data.rds")

# Function to return a forecast of 0s:

forecast.zeros <- function(ts, horizon) {
  # takes a time series and a horizon and returns 
  # a 0 forecast over that horizon
  ts.temp = ts(rep(0,length(ts)), 
                start = start(ts), 
                frequency = frequency(ts))
    ts.fcst = forecast(ts.temp, h=horizon)
    return(ts.fcst)
  }

# create a function to forecast a series based on a ts and df index

output.multiple <- function(isbn){  # note that this should be a 2 param fn
  
  # Generates multiple versions of a forecast to a csv file.  Takes as input 
  # ISBN and implicit time.series and title.data data frames
  
  horizon = 15
  isbn.prev <- previous.edition.isbn(isbn, title.data) # Get previous edition
  ts1 = time.series[, isbn]
  
  num.contig <- length(na.contiguous(ts1))  # how many contiguous values?
  
  if (num.contig > 6){  # if there's enough data, generate a forecast
    
    # base forecast
    base.forecast = forecast(ts1, h=horizon)
    
  } else {   # if there's not enough data, generate a zero forecast
    
    # zeroes forecast
    base.forecast = forecast.zeros(ts1, horizon)
    plot(base.forecast)

  }
  
  plot(base.forecast)
  output <- base.forecast$mean

  # If there's a previous edition, generate additional forecasts

  if (!is.na(isbn.prev)) {
    
    # generate a forecast from current edition plus previous edition
    
    ts.summed <- sum.edition.data(isbn, time.series, title.data)
    summed.prev.fcst = forecast(ts.summed, h = horizon)
    plot(summed.prev.fcst)
    output <- cbind(output, summed.prev.fcst$mean)
    
    # generate a forecast that is simply the previous editions data
    
    prev.edition <- align.previous.edition(isbn, time.series, title.data) # aligns data
    
    # take the window associated with the forecast
    
    prev.edition.2 <- window(prev.edition, 
                      start= start(summed.prev.fcst$mean),
                      end = end(summed.prev.fcst$mean),
                      extend = F)

    output <- cbind(output, prev.edition.2)
    
    }
  
  # print(output)
  
  # note that the date formatting needs to be fixed
  # write.csv(file=paste(isbn, ".csv", sep = ""), x=as.data.frame(output))
  write.zoo(output,file=paste(isbn, ".csv", sep = ""), index.name="Date",sep=",")
}

isbn = "1319056911"  # How Life Works
output.multiple(isbn)

# isbn2 = "0312132921"
# output.multiple(isbn2)
# isbn3 <- "031243829X"
# output.multiple(isbn3)

```


Review Individual Titles
------------------------

```{r}

time.series = readRDS("time_series.rds")
title.data = readRDS("title_data.rds")

title.data$Isbn

isbn1 <- "0312132921"
isbn2 <- "031241501X" # skewed title 1
isbn3 <- "312566735" # skewed title 2
isbn4 <- "145760647X" # skewed title 3

skewed.isbns <- as.list(c(isbn2, isbn3, isbn4))
ts1 = time.series[, isbn4]
sum(!is.na(ts1))
ts1.fcst = forecast(ts1, h = 15)
plot(ts1.fcst)

# test actions.
str(ts1.fcst)
ts1.fcst$method
print(length(na.contiguous(ts1)))


```

Plot two time series on a single axis
-------------------------------------

```{r plot.two}

isbn1 <- "031241501X"
isbn2 <- "031243829X"
time.series = readRDS("time_series.rds")

plot.two <- function(isbn1, isbn2, ts) {
  # Plots two ISBNs in a facet plot, one over the other.  Takes 2 ISBNs as
  # strings and a time series data frome as input and generates two plots.
  plot.data <- ts[,c(isbn1, isbn2)]
  plot.data <- zoo::na.trim(plot.data, sides = "left", is.na = "all")
  plot(plot.data, main = "Comparative Time Series Plot", 
       xlab = "Date",
       ylab = "Units",
       plot.type = "s", 
       col = 1:2)
  legend("topright", 
         legend = c(isbn1, isbn2),
         lty = c(1,1),
         col = 1:2)
  }

plot.two(isbn1, isbn2, time.series)


```

Plot title and previous edition (if any)
----------------------------------------

```{r previous.edition}

time.series = readRDS("time_series.rds")
title.data = readRDS("title_data.rds")

isbn1 = "1319056911"  # How Life Works
isbn2 = "0312132921"
isbn3 <- "031243829X"

plot.previous.edition <- function(isbn, ts.data, title.data){
  # Takes an ISBN as a string and two data frames: title data and title
  # time series as inputs and plots the time series and
  # the previous edition.  If there is no previous edition, it returns 
  # an NA.
  
  isbn.prev <- previous.edition.isbn(isbn, title.data) # Get previous edition
  
  # Fail if there's no previous edition (isbn.prev == NA)
  if (is.na(isbn.prev)) {
    return(NA)
    }
  
  # Fail if previous edition has no data (not yet implemented)
  
  # Plot results
  plot.two(isbn1, isbn.prev, time.series)
}

# Test plot.previous.edition
plot.previous.edition(isbn1, time.series, title.data)
plot.previous.edition(isbn2, time.series, title.data)

isbn3 <- "031243829X"
plot.previous.edition(isbn3, time.series, title.data)

```




Test decomposition with STL

```{r}

isbn = "1319056911"  # How Life Works
isbn.fit = stl(isbn, s.window = "periodic")
plot(isbn.fit)

```

Transplant model from previous edition
--------------------------------------
* Add the previous edition to the current and forecast that
* Decompose the previous edition and use the seasonal factors to forecast the current edition, based on current trend

```{r test.transplant.forecast}

# Make current edition and previous edition start with the same year.

time.series = readRDS("time_series.rds")
title.data = readRDS("title_data.rds")

isbn = "1319056911"  # How Life Works
isbn2 = "0312132921"
isbn3 <- "031243829X"

forecast.with.previous <- function(isbn, ts.data, title.data){
  # Takes an ISBN as a string and two data frames: title data and title
  # time series as inputs and uses the model from the previous edition
  # to forecast the new edition. Returns an NA if there's no previous
  # edition.
  
  isbn.prev <- previous.edition.isbn(isbn, title.data) # Get previous edition
  
  # Fail if there's no previous edition (isbn.prev == NA)
  if (is.na(isbn.prev)) {
    return(NA)
    }
  # Fail if previous edition has no data (not yet implemented)
  ts1 = time.series[, isbn]
  ts1 <- zoo::na.trim(ts1, sides = "both", is.na = "all")
  print(ts1)
  ts1.fcst = forecast(ts1, h=15)
  plot(ts1.fcst)
  
  ts2 = time.series[, isbn.prev]
  ts2 <- zoo::na.trim(ts2, sides = "both", is.na = "all")
  print(ts2)
  ts2.fcst = forecast(ts2, h=15)
  plot(ts2.fcst)
  
  temp = merge.zoo(ts1, ts2)
  temp[is.na(temp)] <- 0
  ts3 = temp[,1] + temp[,2]
  ts3 = ts(ts3, start = start(ts3), frequency = 12)
  ts3[ts3 < 0] <- 0
  print(ts3)
  ts3.fcst = forecast(ts3, h=15)
  plot(ts3.fcst)

}

# Test plot.previous.edition
forecast.with.previous(isbn1, time.series, title.data)
# align.previous.edition(isbn2, time.series, title.data)

isbn3 <- "031243829X"
forecast.with.previous(isbn3, time.series, title.data)  # should fail

```

Recommend titles to examine
---------------------------

```{r}

title.data = readRDS("title_data.rds")
title.forecasts = readRDS("forecasts.rds")

# add another column for whether a review is required

calc.review <- function(forecast, onhand){
  if(forecast == 0){"Insuff data"}
  else if(forecast > onhand){"Print review"}
  else{"No Review"}
}

title.planning <- title.data %>% rowwise() %>% 
  dplyr::mutate(Review = calc.review(title.forecasts$three, (Available + Reserved.Stock + Sets.Made.Up))) %>% 
  dplyr::arrange(Review) %>% 
  mutate(OQ.6 = title.forecasts$six) %>% 
  mutate(OQ.9 = title.forecasts$nine) %>% 
  mutate(OQ.12 = title.forecasts$twelve) %>% 
  mutate(OQ.15 = title.forecasts$fifteen)

# output to csv

saveRDS(title.planning, "title_planning.rds")
write.csv(file="Title_Planning.csv", x=title.planning)

```

Release Notes
-------------
v6
* plot.two function
  * Eliminate leading blanks in plot.two function
  * Moved both lines into same panel for easier comparison
* align.previous edition
  * added function align.previous.edition to produce plot
  * it aligns peaks, not months - this may need to be changed
  * It probablly also needs an export, or to at least to print the prev isbn in the legend
v7
* skipped
v8
* sum.previous.edition function created
* output.multiple function created
